---
title: RIF and PMB combination
output: 
  md_document:
      fig_width: 7
      fig_height: 5
---

# Background 

We looked for combination therapies of neglected and misused antibiotics.
Rifampin and polymyxin B combination stood out as a promising approach against
scpectrum of clinical isolates.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, dpi=300)
knitr::opts_knit$set(global.par = TRUE)
library('data.table')
library('tidyverse')
library('DT')
library('cowplot')
library('BIGL')

devtools::document('input/r')
suppressMessages(devtools::load_all('input/r'))

options(datatable.print.nrows=20)

theme_set(theme_html())
update_geom_defaults("point", list(size = 0.9))
update_geom_defaults("line", list(size = 0.5))
.text_size = 3
.point_size = 2
.facet_size = 8
theme_update(
  #legend.position = "bottom",
  strip.text.y = element_text(size = 8)
)
```

```{r prep}
dat = readRDS('../analysis/data/r_obj/pmb-rif_all-clean.rds')
dat[, effect := round(effect, 3)]

# incomplete data interferes with surface analysis methods
dat = dat[!(cond=='intra' & strain=='ATCC27853' &
  !(c1 %in% c(0, 0.0016, 0.016, 0.048, 0.16, 0.48, 
      4.8, 16, 48, 160, 480, 1600) &
   c2 %in% c(0, 0.0004, 0.003, 0.004, 0.012, 0.03, 
     0.12, 0.3, 1, 1.2, 3, 4, 30, 40, 60, 200)))
]
# arrive at informative and uniform concentration range
dat = dat[(d1 %between% c(0.003, 100) | d1==0) & (d2 %between% c(0.003, 100) | d2==0)]

dat[is.na(effect), effect := -4.5]

# Response surface analysis
# rs = getRS(dat, 'rs')

# # Time-kill ------------------------------
# 
# dtk = fread('input/dat/time-kill.tsv') %>% 
#   remNonCharcoal()
# 
# .myLabels <- c("neg", "RIF 0.3x MIC", "PMB 1x MIC", "combination", 
#   "RIF 3x MIC", "PMB 10x MIC", "PMB 0.3x MIC", "PMB 3x MIC")
# 
# dtk[, new_time_h := ifelse(time_h == 24, 10, time_h)] 
# xaxs = c(0, 2, 4, 6, 10)
# xlabs = c(0, 2, 4, 6, 24)
# dtk[, group := factor(group, levels=1:8, labels=.myLabels)]
```

# Combination

We studied the potential of the rifampicin-polymyxin B combination against
intra- and extracellular forms of bacteria: three _P. aeruginosa_ strains, 
two clinical isolates of _A. baumannii_, _E. cloacae_, and _K. pneumoniae_.


# Time-kill

```{r}

```



# Chemical genetic screen

```{r}

```

