---
title: RIF and PMB combination
output: 
  md_document:
      fig_width: 7
      fig_height: 5
---

# Background 

We looked for combination therapies of neglected and misused antibiotics.
Rifampin and polymyxin B combination stood out as a promising approach against
scpectrum of clinical isolates.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, dpi=300)
knitr::opts_knit$set(global.par = TRUE)
library('data.table')
library('tidyverse')
library('DT')
library('cowplot')
library('BIGL')
library('drc')

devtools::document('input/r')
suppressMessages(devtools::load_all('input/r'))

options(datatable.print.nrows=20)

theme_set(theme_html())
update_geom_defaults("point", list(size = 0.9))
update_geom_defaults("line", list(size = 0.5))
.text_size = 3
.point_size = 2
.facet_size = 8
theme_update(
  #legend.position = "bottom",
  strip.text.y = element_text(size = 8)
)
```

# Combinations

We studied the potential of the rifampicin-polymyxin B combination against
intra- and extracellular forms of bacteria: three _P. aeruginosa_ strains, 
two clinical isolates of _A. baumannii_, _E. cloacae_, and _K. pneumoniae_.


```{r}
dat = readRDS('../analysis/data/r_obj/pmb-rif_all-clean.rds')
dat[, effect := round(effect, 3)]

# incomplete data interferes with surface analysis methods
dat = dat[!(cond=='intra' & strain=='ATCC27853' &
  !(c1 %in% c(0, 0.0016, 0.016, 0.048, 0.16, 0.48, 
      4.8, 16, 48, 160, 480, 1600) &
   c2 %in% c(0, 0.0004, 0.003, 0.004, 0.012, 0.03, 
     0.12, 0.3, 1, 1.2, 3, 4, 30, 40, 60, 200)))
]
# arrive at informative and uniform concentration range
dat = dat[(d1 %between% c(0.003, 100) | d1==0) & (d2 %between% c(0.003, 100) | d2==0)]

dat[is.na(effect), effect := -4.5]

# Response surface analysis
# rs = getRS(dat, 'rs')

```


# Reverse Genetics Screen

To account for the synergy in molecular terms—beyond a nonspecific increase in membrane permeability by polymyxin B—we turned to chemical genetics [Brochado and Typas, 2013](https://doi.org/10.1016/j.mib.2013.01.008). Working with ordered PA14 transposon library [Liberati et al., 2006](https://doi.org/10.1073/pnas.0511100103), we derived a growth measure for monotherapies and combinations using colony opacity [Kritikos et al., 2017](https://doi.org/10.1038/nmicrobiol.2017.14). To account for plate-to-plate variation, the opacity was multiplicatively corrected. This results in zero-centering of the Bliss scores, which were derived next. The significance of difference from zero Bliss score, for any mutant, was estimated by a T-test (5 biological replicates) and corrected for multiple testing (Benjamini-Hochberg).

```{r}
dat = loadIrisFiles("input/dat/raw/iris_files") %>% 
  addIrisVars() %>% 
  # Manual curation by visual inspection of images
  remBadPlates() %>% 
  flagBadColonies() %>% 
  turnBadColoniesToNAs() %>% 
  # Automatic curation by numbers
  setMisdetectedColoniesToNAs() %>%   # colonies missing from 60% of plates
  # plate-to-plate correction, brings plate opacities to same median
  correctOpacityForPlates()

rep.correlations = getPairwiseReplicateCorrelations(dat)
candidate.files.to.kick.out = getCandidateFilesToKickOut(rep.correlations)
candidate.files.to.kick.out[ratio.discordant.vs.all==1, filename]
# [1] "LBpH5.5-Rifampicin16Polymyxin2-015.JPG.iris"

# Fitness ----------------------------------------
condition.vs.control = long2wide(dat)  # also average controls using median
#remember: opacity was replaced by the corrected opacity
condition.vs.control[,f.condition:=opacity/control.median]
dim(condition.vs.control[is.na(f.condition)])[1] / dim(condition.vs.control)[1]
# [1] 0.02289894  #2.2% of fitness values are NA

# Epsilon ----------------------------------------
drug.merge = getEpsilons(condition.vs.control)

drug.merge = addAnnotation(drug.merge)
drug.merge = tTest(drug.merge)
condition.gene.t.test.results = getTTestResults(drug.merge) %>% 
  correctForMultipleTesting()
```

```{r revgen-volc}
tmp = condition.gene.t.test.results %>% 
  mutate(t.test.q.value = ifelse(t.test.q.value < 0.001, 0.001,
      t.test.q.value)) %>% 
  mutate(showname = ifelse(
      (t.test.q.value < 0.05 & gene.name != '') | t.test.q.value < 0.01 ,
      T, F)) 

(p0 = filter(tmp, media == 'LB') %>%  
  {
  ggplot(., aes(epsilon.median.mutant.condition, -log10(t.test.q.value), col = t.test.q.value < 0.05)) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c('grey70', 'grey20')) +
  ggrepel::geom_text_repel(data= . %>% .[(showname)], size = 6,
    aes(label=gene.name.to.show), box.padding=0.75, seed=3) +
  lims(x = c(-0.8, 0.6)) +
  labs(x = 'Shift in Bliss score', y = expression('log'[10]*'(p-value)')) +
  theme(aspect.ratio = 1, legend.position = 'none')

})

# ggsave('output/fig/ChemGen_LB.svg', width=5, height=5)

p0 %+% filter(tmp, media == 'LBpH5.5')
# ggsave('output/fig/ChemGen_LBpH5.5.svg', width=5, height=5)
```

```{r revgen-venn}
getGenes = function (dat) {
  out = list()
  for(i in unique(dat$media))
    out[[i]] = unique(dat[t.test.q.value < 0.05 & media == i], by='locus') %>%
      .[, (gene.name.to.show)]
  return(out)
}

foo = getGenes(condition.gene.t.test.results)
vtable = gplots::venn(foo)
```

# Screen Validation

We validate the sensitivity of identified candidate mutants in low throughput
and in liquid medium. Fixed ratio design.

# Supplementary Figures

## Dose Response And pH

```{r}
theme_set(theme_pdf())

sub = subset(dat, strain == 'ATCC27853' & (d1 == 0 | d2 == 0))

p1 = function() pltDR(effect ~ c1, subset(sub, c2==0), xlab = 'RIF, mg/L', zero=30)
p2 = function() pltDR(effect ~ c2, subset(sub, c1==0), col='orange', xlab = 'PMB, mg/L')

# pH
foo = fread("input/dat/raw/growth_and_pH.tsv")

foo = foo[time_h >= 0 & overnight_medium == 'pH7.4', .(
  od = gmu(od620),
  od.ci = gci(od620),
  ph = mean(pH),
  ph.ci = ci(pH)
  ), .(growth_medium, time_h)]


p0 = foo %>% 
  mutate(
    growth_medium = relevel(factor(growth_medium), 'pH7.4'),
    time_h = ifelse(time_h > 20, 20, time_h),
    NULL
    ) %>% 
  ggplot(aes(time_h, od, shape = growth_medium)) +
  scale_x_continuous(breaks = c(0, 5, 10, 20), labels = c(seq(0, 10, 5), 24), limits = c(0, 20)) +
  scale_shape_manual(name = '', values = c(19, 22, 15), labels = c('pH 7.4', 'pH 5.5',
      'pH 5.5 buffered')) +
  theme(legend.position = 'top')

p3 = p0 + 
  scale_y_continuous(trans='log2', breaks = c(0.001, 0.01, 0.1, 1), limits = c(0.001, 10)) + 
  geom_point(size=.point_size, col='grey20') +
  geom_pointrange(aes(ymin = od*od.ci, ymax = od/od.ci)) +
  geom_line(col='grey20') +
  labs(x = 'Time, h', y = expression('OD'[620]))

p4 = p0 %+% aes(y = ph) + 
  scale_y_continuous(position = 'right') +
  geom_point(size=.point_size, col=cbPalette[3]) +
  geom_line(col=cbPalette[3]) +
  geom_pointrange(aes(ymin = ph-ph.ci, ymax = ph + ph.ci), col=cbPalette[3]) +
  theme(legend.position = 'none',
    axis.title.y.right=element_text(color = cbPalette[3]),
    axis.text.y.right=element_text(color = cbPalette[3]),
    axis.ticks.y.right=element_line(color = cbPalette[3])
  ) +
  labs(x = '', y = 'pH')

aligned = align_plots(p3, p4, align = 'hv', axis = 'tblr')

plot_grid(
  ggdraw(p1), ggdraw(p2), ggdraw(aligned[[1]]) + draw_plot(aligned[[2]]), labels = 'AUTO', align = 'v'
)

# ggsave('output/fig/SFig_monotherapy_dose-response_and_pH.png', width=10, height=9)
```


## Time-kill

```{r}
# # Time-kill ------------------------------
# 
# dtk = fread('input/dat/time-kill.tsv') %>% 
#   remNonCharcoal()
# 
# .myLabels = c("neg", "RIF 0.3x MIC", "PMB 1x MIC", "combination", 
#   "RIF 3x MIC", "PMB 10x MIC", "PMB 0.3x MIC", "PMB 3x MIC")
# 
# dtk[, new_time_h := ifelse(time_h == 24, 10, time_h)] 
# xaxs = c(0, 2, 4, 6, 10)
# xlabs = c(0, 2, 4, 6, 24)
# dtk[, group := factor(group, levels=1:8, labels=.myLabels)]

```
