---
title: RIF and PMB combination
output: 
  md_document:
      fig_width: 7
      fig_height: 5
---

# Background 

We looked for combination therapies of neglected and misused antibiotics.
Rifampin and polymyxin B combination stood out as a promising approach against
scpectrum of clinical isolates.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, dpi=300)
knitr::opts_knit$set(global.par = TRUE)
library('data.table')
library('tidyverse')
library('DT')
library('cowplot')
library('BIGL')

devtools::document('input/r')
suppressMessages(devtools::load_all('input/r'))

options(datatable.print.nrows=20)

theme_set(theme_html())
update_geom_defaults("point", list(size = 0.9))
update_geom_defaults("line", list(size = 0.5))
.text_size = 3
.point_size = 2
.facet_size = 8
theme_update(
  #legend.position = "bottom",
  strip.text.y = element_text(size = 8)
)
```

# Combination

We studied the potential of the rifampicin-polymyxin B combination against
intra- and extracellular forms of bacteria. Three _P. aeruginosa_ strains and
two clinical isolates of A. baumannii, E. cloacae, and K. pneumoniae were
investigated.

```{r}

```


# Time-kill



# Chemical genetic screen
